---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# simloglm

<!-- badges: start -->
<!-- badges: end -->

The goal of simloglm is to provide functions to simulate the correct quantities of interest from linear regression models with logged dependent variables.

This package implements the method of Rittmann, Neunhoeffer and Gschwend (2025), "How to improve the substantive interpretation of regression results when the dependent variable is logged", *Political Science Research and Methods* 13(1), 203-211, [doi:10.1017/psrm.2023.29](https://doi.org/10.1017/psrm.2023.29). Please cite the paper if you use the package; `citation("simloglm")` gives both entries.

## Installation

Currently you can install the development version of simloglm from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mneunhoe/simloglm")
```
## Usage

```{r example}
library(simloglm)
df <- cars
regression <- lm(log(dist) ~ speed, data = df)

# Specifying no scenario simulates at the mean of every variable.
simulation_results_average <- simloglm(regression, verbose = FALSE)

# Explicitly specifying a scenario.
simulation_results_scenario <- simloglm(regression,
                                        scenario = list(speed = c(5, 10, 20)),
                                        verbose = FALSE)

get_summary(simulation_results_scenario, which_qoi = "median")
get_summary(simulation_results_scenario, which_qoi = "mean")
```

`median` holds the conditional geometric mean (which is also the conditional
median) of the dependent variable, `mean` the conditional arithmetic mean, each
as a matrix with one row per simulation draw and one column per scenario.

Scenario values are given on the **original scale** of the variables, so a
transformation in the formula is applied for you:

```{r loglog}
loglog <- lm(log(dist) ~ log(speed), data = df)
res <- simloglm(loglog, scenario = list(speed = c(10, 11)), verbose = FALSE)

# A ten percent increase in speed multiplies the median by 1.1^beta.
get_ratio(res)$ratio
```

Variables you leave out of the scenario are held at their mean (numeric) or
reference level (factor). See `vignette("log_log_models")` for elasticities and
first differences, and the other vignettes for replications of published work.

## Issues
Please reach out in case you find bugs and open an issue on Github. 
