---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# simloglm

<!-- badges: start -->
<!-- badges: end -->

The goal of simloglm is to provide functions to simulate the correct quantities of interest from linear regression models with logged dependent variables. This paper accompanies the paper "How to Improve the Substantive Interpretation of
Regression Results when the Dependent Variable is
logged".

## Installation

Currently you can install the development version of simloglm from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("mneunhoe/simloglm")
```
## Usage

This is a basic example which shows you how to solve a common problem:

```{r example}
library(simloglm)
df <- cars
regression <- lm(log(dist)~speed, data = df)
# Specifiying no scenario to simulate at the mean of speed.
simulation_results_average <- simloglm(regression)
# Explicitily specifying a scenario.
simulation_results_scenario <- simloglm(regression, scenario = list(speed = c(5, 10, 20)))
```

See the vignettes for replications of published work.

## Issues
Please reach out in case you find bugs and open an issue on Github. 
