---
title: "Log-log models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Log-log models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(simloglm)
set.seed(20220609)
```

A log-log model regresses `log(y)` on `log(x)`, so its slope is an elasticity:
a one percent increase in `x` is associated with a `beta` percent change in `y`.
This vignette shows how to read that off a `simloglm` object, and points out one
thing about the sigma^2 / 2 correction that surprises people.

# Scenarios are given on the original scale

```{r}
fit <- lm(log(dist) ~ log(speed), data = cars)
coef(fit)
```

You supply `speed`, not `log(speed)`. The `log()` in the formula is applied for
you, exactly as it is by `predict()`:

```{r}
res <- simloglm(fit, scenario = list(speed = c(10, 11)), verbose = FALSE)
get_summary(res, which_qoi = "median")
```

The one place this does not hold is `X`: a model matrix passed directly is used
as-is, so its columns must already be transformed.

A caveat on the default scenario. With no `scenario`, `simloglm()` holds every
numeric variable at its arithmetic mean *on the original scale*. In a log-log
model "x at its mean" is ambiguous, because the mean of `log(x)` corresponds to
the **geometric** mean of `x`, which is smaller. If you want the geometric mean,
ask for it:

```{r}
mean(cars$speed)
exp(mean(log(cars$speed)))

simloglm(fit, scenario = list(speed = exp(mean(log(cars$speed)))),
         verbose = FALSE)$median_point_estimate
```

# Recovering 1.1^beta from a ratio

The ratio of the two scenarios above is the multiplicative effect of a ten
percent increase in `speed`, and it should be `1.1^beta`:

```{r}
ratio <- get_ratio(res)
ratio$ratio
1.1^coef(fit)[2]
ratio$quantiles
```

So a ten percent higher speed goes with roughly a
`r round(100 * (get_ratio(res)$ratio - 1))` percent higher median stopping
distance.

# The sigma^2 / 2 correction cancels in a ratio

The point of this package is the correction that turns a prediction on the log
scale into the conditional *arithmetic* mean on the original scale. For a ratio
it makes no difference at all:

$$\frac{\exp(X_2\beta + \sigma^2/2)}{\exp(X_1\beta + \sigma^2/2)} = \exp((X_2 - X_1)\beta).$$

In the fast path this cancels draw by draw, because row `i` of both scenarios
uses the same `sigma2_sim[i]`. The two quantities of interest are therefore
numerically identical:

```{r}
all.equal(get_ratio(res, which_qoi = "median")$ratio,
          get_ratio(res, which_qoi = "mean")$ratio)

all.equal(res$median[, 2] / res$median[, 1],
          res$mean[, 2] / res$mean[, 1])
```


# Where the correction matters

Level predictions and first differences. A first difference is measured in units
of `y`, so it depends on the baseline you start from, and the correction shifts
that baseline:

```{r}
res3 <- simloglm(fit, scenario = list(speed = c(10, 11, 20)), verbose = FALSE)

fd_median <- get_first_difference(res3, which_qoi = "median")
fd_mean <- get_first_difference(res3, which_qoi = "mean")

c(median = fd_median$fd, mean = fd_mean$fd)
```

Always report a first difference together with its baseline, which is why
`get_first_difference()` returns one:

```{r}
data.frame(
  qoi = c("median", "mean"),
  baseline = c(fd_median$baseline, fd_mean$baseline),
  first_difference = c(fd_median$fd, fd_mean$fd),
  lower = c(fd_median$quantiles[1], fd_mean$quantiles[1]),
  upper = c(fd_median$quantiles[2], fd_mean$quantiles[2])
)
```

The same ten percent increase produces a different absolute difference depending
on where it starts:

```{r}
c(from_10_to_11 = get_first_difference(res3, which_scenarios = c(1, 2))$fd,
  from_10_to_20 = get_first_difference(res3, which_scenarios = c(1, 3))$fd)
```

# A proportional counterfactual on the observed sample

Setting every observation to the same value of `speed` is a strange
counterfactual under an elasticity. `multiplier` shifts each observation
proportionally instead, within the observed value approach:

```{r}
ov <- simloglm(fit,
               multiplier = list(speed = c(1, 1.1)),
               observed_value_approach = TRUE,
               verbose = FALSE)

get_summary(ov)
get_ratio(ov)$ratio
```

Because `log(1.1 * x) = log(1.1) + log(x)`, a proportional shift of a logged
variable is a constant shift on the log scale, so this ratio again recovers
`1.1^beta`, averaged over the observed sample.
